Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-10-20 paper/main.tex
3.1

\section{Artist} 
In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations. We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}. The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$.

\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
  \label{eq:artist:artist}
\end{equation}

\note{roadmap the rest of this section}

@@

The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves. The category $\mathcal{S}$ is the category of open sets $\{U\}$ over the graphic base space \gbase. The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}. A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space; bar plots and scatter plots on a 2D screen are two such objects.

As will be discussed in \autoref{sec:artist:construction}, the  $\vartist$ is constructed in such a way that it generates a single type of graphic;

Therefore, we define the object as a morphism from data to the sheaf mage of the artist in the graphic space $\vartist: \mathcal{\dtotal} \rightarrow Im_A \subset \mathcal{H}$

\begin{equation}
  \label{eq:artist:morphism-of-sheaf}
  \begin{tikzcd}
    & \mathcal{T} & & & \mathcal{P} & \\
{} \arrow[rr, "f", Rightarrow, shift right, shorten=3em] & 
& 
{} \arrow[r, "\vartist", shorten=-1.25em, bold] & {} \arrow[rr, "g", Rightarrow, shift right, shorten=3em] & & {} \\
& \mathcal{\dbase}^{op} \arrow[uu, "\mathcal{\dtotal}", bend left] 
  \arrow[uu, "\mathcal{\dtotal}"', bend right] 
  & & 
& \mathcal{\gbase}^{op} \arrow[uu, "\mathcal{\gtotal}", bend left] 
  \arrow[uu, "\mathcal{\gtotal}"', bend right] &   
  \end{tikzcd}
\end{equation}

% f is a natural transform, all transforms on the sheafs commute
% g is a natural transform ...
% space of g->g is too big, so we look at Im...
% b/c it's a natural transform, - f carries i over b/c it's a natural transformation

therefore the artist is a map between functors
\begin{equation}
  \tag{\ref{eq:artist:artist}}
  \vartist:\mathcal{\dtotal}\rightarrow{\mathcal{\gtotal}} 
\end{equation} 
which means that the artist is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}.

% such that (9)

Extra property that implies equivariance is preserved is illustrated in the diagram which must commute
%% note Im_A in \H, A_M (lifting/pushforward)
\begin{equation}
  \label{eq:artist:equivariance}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "{Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \in \mathcal{\gtotal} \arrow[d, "\vartist_{Hom_{\mathcal{\dfiber}}}"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist} \in \mathcal{\gtotal}                                                
  \end{tikzcd}
\end{equation}

such that the morphisms on the graphic category $\mathcal{H}$ are homomorphic to the morphisms $\monoid$ on the fibers $\mathcal{F}$ \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. 
%% such that monoid on this side is monoid on that side

In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity. We then use this formulation to guide the development of visualization 
library components in \autoref{sec:case-study}.

>>>>

\section{Artist} 
In this section, we use category theory to formally express the implicit assumptions that visualization library components make in transforming data into graphical representations. We propose that the visualization transformation can be modeled as a functor, which we call the \textcolor{artist}{Artist} $\mathcal{\vartist}$\footnote{We call this transformation the artist because the \mintinline{python}{Artist} object in Matplotlib\cite{hunterMatplotlib2DGraphics2007}}. The artist \vartist\ converts data, which we model as the catgory $\mathcal{\dtotal}$ to graphics, which we model as the category $\mathcal{\gtotal}$. We formulate this association as

\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow \mathcal{\gtotal}
  \label{eq:artist:artist}
\end{equation}
In this section we describe the structure of the data and graphic spaces, which we formulate as objects in categories and the morphisms between those objects. We then use these definitions to express the structure that an artist must preserve. Finally, we propose that artists can be composed by manufacturing new artists based on various types of input data.

@@

The graphic category $\mathcal{\gtotal}$ consists of the functor object $\sheaf_{\mathcal{\gtotal}}: \mathcal{S}^{op} \rightarrow \Gamma(\mathcal{S}, \gtotal)$ and morphisms between sheaves. The category $\mathcal{S}$ is the category of open sets $\{U\}$ over the graphic base space \gbase. The set of sets of sections $\Gamma(\mathcal{S}, \gtotal)$ is a subcatagory of \texbf{Set}. A set $\Gamma(U_i, \gtotal)$ contains functions to generate arbitrary graphics with a shared continuity and display space, while the artist $\vartist$ is constructed such that it generates a single type of graphic, as described in \autoref{sec:artist:construction}. Therefore, the range of the the artist functor, which we denote $Im_{\vartist}$, is a sheaf map $Im_{\vartist}$. The sheaf $Im_{\vartist}$ is a map between the category of open sets $\gtotal^{op}$ and a subcategory of \textbf{Set}. In this subcatagory, each set of sections in the sheaf $Im_{\vartist}$ over an arbitrary open set $U_{i}$ is a subset of the corresponding full set of sections $\Gamma(U_i, \gtotal)$.

Since the artist is a morphism of sheafs
\begin{equation}
  \vartist: \mathcal{\dtotal} \rightarrow Im_A
\end{equation}
it is a natural transformation\cite{SheafMathematics2021,bradleyWhatNaturalTransformation}. This means that, by definition, all morphisms $\varphi$ on $\mathcal{\dtotal}$ are expected to commute \cite{spanier1989algebraic,fongInvitationAppliedCategory2019}. This is the basis of the equivariance illustrated in \autoref{eq:artist:equivariance}

\begin{equation}
  \label{eq:artist:equivariance}
  \begin{tikzcd}
    \mathcal{\dtotal} \arrow[r, "\vartist"] \arrow[d, "\varphi \in {Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)}"'] & Im_{\vartist} \arrow[d, "\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}"] \\
    \mathcal{\dtotal} \arrow[r, "\vartist"]                                                           & Im_{\vartist}                                                 
  \end{tikzcd}
\end{equation}
which states that transforms on the data side $\varphi \in Hom_{\mathcal{\dfiber}}(\dfiber, \dfiber)$ have an approximately equivalent transform in the data side $\vartist_{\varphi \in Hom_{\mathcal{\dfiber}}}$. This commutativity is described as the matching of data and visual structure by Kindlmann and Scheidegger when introducing their analougous correspondence principal. In \autoref{sec:artist:construction}, we show a method of constructing the artist such that the constraints of \autoref{eq:artist:morphism-of-sheaf} and \autoref{eq:artist:equivariance} can be satisfied, thereby ensuring equivariance and the preservation of continuity. We then use this formulation to guide the development of visualization library components in \autoref{sec:case-study}.
